Preferences for the commodification of pensions in Chile: the rol of intergenerational social mobility

### 2025-11-03 paper/07-supplementary-material.qmd
Up SP

---
title: "Supplementary material"
format:
  html: 
    code-fold: true
    html-math-method: katex
    number-sections: true
  docx:
    number-sections: true
  pdf: 
    geometry: margin=0.5cm
    template-partials: 
      - title.tex
    keep-tex: true
    number-sections: true
editor: source
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
colorlinks: true
fig-height: 4
fig-width: 7.5
editor_options: 
  chunk_output_type: console
---

@@

```{r}
#| label: fig-interact
#| fig-cap: "Balance SMD — Mobility treatment (ATT)"
#| fig-cap-location: bottom
#| echo: false
#| results: asis
#| warning: false
#| out-width: '90%'
#| fig-asp: 0.9

>>>>

# Supplementary material {#anexo}

@@

```{r}
#| label: fig-balance
#| fig-cap: "Balance SMD — Mobility treatment (ATT)"
#| fig-cap-location: bottom
#| echo: false
#| results: asis
#| warning: false
#| out-width: '90%'
#| fig-asp: 0.9

### 2025-11-03 paper/paper.tex
Up SP

\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Preferences for the commodification of pensions in Chile: the rol of intergenerational social mobility},
  pdfauthor={Andreas Laffert Tamayo},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\setstretch{1.5}
This document was last modified at 2025-11-03 17:04:08 and it was last
rendered at 2025-11-03 17:04:08.

@@

>>>>

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Preferences for the commodification of pensions in Chile: the rol of intergenerational social mobility},
  pdfauthor={Andreas Laffert Tamayo},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\setstretch{1.5}
This document was last modified at 2025-11-03 17:13:25 and it was last
rendered at 2025-11-03 17:13:25.

@@

\section{Supplementary material}\label{anexo}

This section presents the supplementary material for this study. First,
graphs are shown for the evaluation of the balance of IPW-ATT via
entropy balance. Second, the complete estimated regression models are
shown. Finally, the sensitivity analyses performed are shown.

\includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-balance-1.pdf}

\caption{\label{fig-balance}Balance SMD --- Mobility treatment (ATT)}

\caption{\label{tbl-complete1}Effects of intergenerational occupational
mobility on preferences for pension commodification, with covariates and
wave fixed effects.}

\begin{center}
\scalebox{0.7}{
\begin{tabular}{l c c c c c c}
\hline
 & Low-Middle & Low-High & Middle-High & Middle-Low & High-Middle & High-Low \\
\hline
Intercept                                               & $0.18$            & $-0.02$           & $0.30^{*}$        & $0.23$            & $0.13$            & $0.14$            \\
                                                        & $ [-0.36;  0.71]$ & $ [-0.35;  0.31]$ & $ [ 0.01;  0.58]$ & $ [-0.08;  0.54]$ & $ [-0.19;  0.46]$ & $ [-0.28;  0.56]$ \\
Mobility treatment                                      & $-0.02$           & $-0.02$           & $0.09^{*}$        & $0.00$            & $-0.07^{*}$       & $-0.10^{*}$       \\
                                                        & $ [-0.08;  0.05]$ & $ [-0.09;  0.05]$ & $ [ 0.02;  0.16]$ & $ [-0.06;  0.06]$ & $ [-0.15; -0.00]$ & $ [-0.18; -0.02]$ \\
Age (in years)                                          & $0.00$            & $0.00$            & $-0.00$           & $-0.00$           & $0.00$            & $0.00$            \\
                                                        & $ [-0.00;  0.00]$ & $ [-0.00;  0.00]$ & $ [-0.00;  0.00]$ & $ [-0.00;  0.00]$ & $ [-0.00;  0.01]$ & $ [-0.00;  0.01]$ \\
Female (Ref. = Male)                                    & $-0.14^{*}$       & $-0.14^{*}$       & $-0.13^{*}$       & $-0.07^{*}$       & $-0.11^{*}$       & $-0.12^{*}$       \\
                                                        & $ [-0.20; -0.08]$ & $ [-0.22; -0.06]$ & $ [-0.20; -0.06]$ & $ [-0.14; -0.01]$ & $ [-0.19; -0.03]$ & $ [-0.21; -0.03]$ \\
Chilean nacionality (Ref. = Non-Chilean)                & $-0.03$           & $0.20$            & $0.08$            & $0.04$            & $-0.00$           & $0.06$            \\
                                                        & $ [-0.63;  0.57]$ & $ [-0.05;  0.46]$ & $ [-0.19;  0.36]$ & $ [-0.26;  0.34]$ & $ [-0.27;  0.27]$ & $ [-0.30;  0.42]$ \\
Indigenous ethnicity (Ref. = Non-indigenous)            & $0.03$            & $-0.05$           & $0.05$            & $-0.01$           & $0.07$            & $0.01$            \\
                                                        & $ [-0.06;  0.12]$ & $ [-0.15;  0.05]$ & $ [-0.08;  0.18]$ & $ [-0.10;  0.08]$ & $ [-0.08;  0.22]$ & $ [-0.13;  0.14]$ \\
Co-residence with both parents (Ref. = No co-residence) & $-0.00$           & $0.04$            & $-0.06$           & $-0.01$           & $0.10^{*}$        & $0.10^{*}$        \\
                                                        & $ [-0.08;  0.07]$ & $ [-0.04;  0.12]$ & $ [-0.14;  0.01]$ & $ [-0.09;  0.06]$ & $ [ 0.01;  0.19]$ & $ [ 0.01;  0.19]$ \\
Parental education                                      & $0.02$            & $0.01$            & $-0.02^{*}$       & $-0.01$           & $-0.00$           & $-0.01$           \\
                                                        & $ [-0.00;  0.05]$ & $ [-0.01;  0.04]$ & $ [-0.03; -0.01]$ & $ [-0.02;  0.01]$ & $ [-0.02;  0.02]$ & $ [-0.03;  0.02]$ \\
Wave (Ref.= 2016)                                       &                   &                   &                   &                   &                   &                   \\
                                                        &                   &                   &                   &                   &                   &                   \\
\quad Wave 2018                                         & $0.02$            & $-0.01$           & $0.08^{*}$        & $0.04$            & $-0.00$           & $-0.00$           \\
                                                        & $ [-0.05;  0.09]$ & $ [-0.09;  0.07]$ & $ [ 0.02;  0.14]$ & $ [-0.02;  0.10]$ & $ [-0.08;  0.08]$ & $ [-0.09;  0.09]$ \\
\quad Wave 2023                                         & $0.06$            & $0.04$            & $0.15^{*}$        & $0.14^{*}$        & $0.12^{*}$        & $0.10$            \\
                                                        & $ [-0.02;  0.13]$ & $ [-0.06;  0.13]$ & $ [ 0.08;  0.22]$ & $ [ 0.07;  0.21]$ & $ [ 0.02;  0.21]$ & $ [-0.00;  0.21]$ \\
\hline
R$^2$                                                   & $0.04$            & $0.04$            & $0.07$            & $0.03$            & $0.05$            & $0.06$            \\
Adj. R$^2$                                              & $0.03$            & $0.03$            & $0.06$            & $0.02$            & $0.04$            & $0.05$            \\
Num. obs.                                               & $878$             & $745$             & $716$             & $773$             & $912$             & $861$             \\
RMSE                                                    & $0.40$            & $0.37$            & $0.39$            & $0.39$            & $0.44$            & $0.42$            \\
N Clusters                                              & $497$             & $449$             & $443$             & $470$             & $536$             & $525$             \\
\hline
\multicolumn{7}{l}{\scriptsize{Note: Cells contain regression coefficients with confidence intervals in parentheses. $^*$ Null hypothesis value outside the confidence interval..}}
\end{tabular}
\label{table:coefficients}
\end{center}

\caption{\label{tbl-complete2}Effects of intergenerational occupational
mobility on preferences for pension commodification.}

\begin{center}
\scalebox{0.8}{
\begin{tabular}{l c c c c c c}
\hline
 & Low-Middle & Low-High & Middle-High & Middle-Low & High-Middle & High-Low \\
\hline
Intercept          & $0.21^{*}$       & $0.18^{*}$       & $0.17^{*}$      & $0.19^{*}$       & $0.32^{*}$       & $0.28^{*}$        \\
                   & $ [ 0.16; 0.26]$ & $ [ 0.13; 0.23]$ & $ [0.12; 0.21]$ & $ [ 0.15; 0.24]$ & $ [ 0.26; 0.38]$ & $ [ 0.21;  0.35]$ \\
Mobility treatment & $-0.01$          & $-0.02$          & $0.08^{*}$      & $-0.00$          & $-0.07$          & $-0.09^{*}$       \\
                   & $ [-0.08; 0.05]$ & $ [-0.08; 0.05]$ & $ [0.01; 0.15]$ & $ [-0.06; 0.06]$ & $ [-0.15; 0.00]$ & $ [-0.17; -0.02]$ \\
\hline
R$^2$              & $0.00$           & $0.00$           & $0.01$          & $0.00$           & $0.01$           & $0.01$            \\
Adj. R$^2$         & $-0.00$          & $-0.00$          & $0.01$          & $-0.00$          & $0.01$           & $0.01$            \\
Num. obs.          & $878$            & $745$            & $716$           & $773$            & $912$            & $861$             \\
RMSE               & $0.40$           & $0.38$           & $0.40$          & $0.39$           & $0.45$           & $0.43$            \\
N Clusters         & $497$            & $449$            & $443$           & $470$            & $536$            & $525$             \\
\hline
\multicolumn{7}{l}{\scriptsize{Note: Cells contain regression coefficients with confidence intervals in parentheses. $^*$ Null hypothesis value outside the confidence interval..}}
\end{tabular}
}
\label{table:coefficients}
\end{center}

\caption{\label{tbl-complete3}Interactions effects between
intergenerational occupational mobility and perceived meritocracy on
preferences for pension commodification, with covariates and wave fixed
effects.}

\begin{center}
\scalebox{0.7}{
\begin{tabular}{l c c c c c c}
\hline
 & Low-Middle & Low-High & Middle-High & Middle-Low & High-Middle & High-Low \\
\hline
Intercept                                                    & $0.15$           & $-0.04$          & $0.30^{*}$       & $0.21$           & $0.08$           & $0.05$           \\
                                                             & $ [-0.36; 0.67]$ & $ [-0.40; 0.31]$ & $ [ 0.04; 0.56]$ & $ [-0.09; 0.51]$ & $ [-0.25; 0.40]$ & $ [-0.30; 0.41]$ \\
Mobility treatment                                           & $-0.02$          & $-0.03$          & $0.07$           & $-0.00$          & $-0.07$          & $-0.08$          \\
                                                             & $ [-0.09; 0.04]$ & $ [-0.10; 0.05]$ & $ [-0.01; 0.14]$ & $ [-0.07; 0.06]$ & $ [-0.15; 0.00]$ & $ [-0.15; 0.00]$ \\
High meritocracy perception (Ref.= Low)                      & $0.05$           & $0.06$           & $-0.01$          & $0.03$           & $0.17^{*}$       & $0.26^{*}$       \\
                                                             & $ [-0.06; 0.16]$ & $ [-0.07; 0.19]$ & $ [-0.09; 0.08]$ & $ [-0.07; 0.13]$ & $ [ 0.00; 0.33]$ & $ [ 0.04; 0.48]$ \\
Mobility treatment x High meritocracy perception (Ref.= Low) & $0.04$           & $0.05$           & $0.15$           & $0.04$           & $-0.03$          & $-0.14$          \\
                                                             & $ [-0.12; 0.20]$ & $ [-0.14; 0.24]$ & $ [-0.03; 0.33]$ & $ [-0.11; 0.19]$ & $ [-0.23; 0.17]$ & $ [-0.38; 0.10]$ \\
\hline
Controls                                                     & Yes              & Yes              & Yes              & Yes              & Yes              & Yes              \\
R$^2$                                                        & $0.04$           & $0.05$           & $0.08$           & $0.04$           & $0.07$           & $0.10$           \\
Adj. R$^2$                                                   & $0.03$           & $0.04$           & $0.07$           & $0.02$           & $0.06$           & $0.08$           \\
Num. obs.                                                    & $878$            & $745$            & $716$            & $773$            & $912$            & $861$            \\
RMSE                                                         & $0.40$           & $0.37$           & $0.39$           & $0.39$           & $0.44$           & $0.41$           \\
N Clusters                                                   & $497$            & $449$            & $443$            & $470$            & $536$            & $525$            \\
\hline
\multicolumn{7}{l}{\scriptsize{Note: Cells contain regression coefficients with confidence intervals in parentheses. $^*$ Null hypothesis value outside the confidence interval..}}
\end{tabular}
}
\label{table:coefficients}
\end{center}

\caption{\label{tbl-holm}Holm correction}

\centering\centering
\begin{tabular}{lll}
\toprule
\textbf{Mobility treatment} & \textbf{p value} & \textbf{Holm}\\
\midrule
Low→Middle & 0.616 & 1.000\\
Low→High & 0.594 & 1.000\\
Middle→Low & 0.904 & 1.000\\
Middle→High & 0.012 & 0.070\\
High→Middle & 0.048 & 0.192\\
\addlinespace
High→Low & 0.017 & 0.083\\
\bottomrule
\multicolumn{3}{l}{\textsuperscript{} Note: p values derived from Table 1}\\
\end{tabular}

\caption{\label{tbl-evalues}E-values for ATT estimates by mobility
trajectory (Ding \& VanderWeele, 2016 approximation)}

\centering\centering
\fontsize{8}{10}\selectfont
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\toprule
\textbf{Mobility treatment} & \textbf{Beta} & \textbf{CI 95\% [low-upper]} & \textbf{sdY} & \textbf{RR point} & \textbf{RR point upper} & \textbf{E point} & \textbf{RR CI} & \textbf{RR CI Upper} & \textbf{E CI} & \textbf{CI Crosses 0}\\
\midrule
Low→Middle & -0.016 & -0.078; 0.046 & 0.403 & 0.965 & 1.036 & 1.230 & 1.110 & 1.110 & 1.000 & 1\\
Low→High & -0.019 & -0.088; 0.051 & 0.383 & 0.956 & 1.046 & 1.264 & 1.128 & 1.128 & 1.000 & 1\\
Middle→Low & 0.004 & -0.056; 0.063 & 0.392 & 1.008 & 1.008 & 1.101 & 0.879 & 1.138 & 1.000 & 1\\
Middle→High & 0.088 & 0.020; 0.156 & 0.403 & 1.219 & 1.219 & 1.736 & 1.046 & 1.046 & 1.264 & 0\\
High→Middle & -0.074 & -0.148;-0.001 & 0.455 & 0.862 & 1.160 & 1.590 & 0.999 & 1.001 & 1.037 & 0\\
\addlinespace
High→Low & -0.098 & -0.177;-0.018 & 0.431 & 0.814 & 1.229 & 1.759 & 0.963 & 1.039 & 1.239 & 0\\
\bottomrule
\end{tabu}
